---
title: "Problem Set 4"
author: "Caleb, Jonathan, Levi"
date: "Tuesday, February 23, 2016"
output: pdf_document
---

# Problem 15

You take numsim samples of size n from a gamma distribution with parameters $\hat{\alpha}$ and $\hat{\lambda}$ and find the upper quartile of each sample. You use the standard deviation of these numsim samples as the estimation of the standard error of $q_hat$.

```{r}
library(dplyr)
library(mosaic)

data60 <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/Illinois60.csv")
names(data60)[1] <- "rain"
data61 <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/Illinois61.csv")
names(data61)[1] <- "rain"
data62 <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/Illinois62.csv")
names(data62)[1] <- "rain"
data63 <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/Illinois63.csv")
names(data63)[1] <- "rain"
data64 <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/Illinois64.csv")
names(data64)[1] <- "rain"

rain <- rbind(data60, data61, data62, data63, data64)
```

```{r}
numsim <- 10000

findQuant <- function(quantile){
  x <- sample(rain$rain, size = length(rain$rain), replace = TRUE)
  return(quantile(x, quantile)[[1]])
}

quants <- do(numsim)*findQuant(.75) # Generating observations of sample quantiles
sd(quants$findQuant)

histogram(~findQuant, data=quants, breaks = 15, type = "density", fit = "normal")

quantile(rain$rain, .75)
confint(quants$findQuant, level = .95)
```

```{r, fig.keep='last'}
xbar <- mean(rain$rain)
sigma <- sd(rain$rain)*221/222 #comment, etc.
qgamma(.75, shape = xbar^2/sigma^2, rate = xbar/sigma^2)
hist(rain$rain, prob=TRUE, main = "Histogram of Rain", xlab = "Inches of Rain")
curve(dgamma(x, shape = xbar^2/sigma^2, rate = xbar/sigma^2), col = "blue", add = TRUE)
abline(v = qgamma(.75, shape = xbar^2/sigma^2, rate = xbar/sigma^2), col = "red")
```


# Problem 17

## Part (a)

Plot various density curves for different values of alpha.

```{r, fig.keep='last'}
curve(dbeta(x, 6, 6))
curve(dbeta(x, 4, 4), add=TRUE)
curve(dbeta(x, 2, 2), add=TRUE)
```


## Part (b)

Sample from the distribution (with some given alpha), find $\mu_2$, and plug into derived formula to get $\hat{\alpha}$

## Part (c)

## Part (d)

Asymptotic variance formula given on page 277.

## Part (e)

Sufficient statistic formula given on page 306.