---
title: 'Problem Set #3 - Florence'
author: "Azka, Jonathan, Jordan"
date: "Monday, February 15, 2016"
output: pdf_document
---

```{r, include=FALSE}
library(mosaic)
```

# Problem 10.6
Various chemical tests were conducted on beeswax. In particular, the percentage of hydrocarbons in each sample of wax was determined.

## Part (a)
Plot the ecdf, the histogram, and a normal probability plot of the percentages of hydrocarbons. Find the .90, .75, .50, .25, and .10 quantiles. Does the distribution appear Gaussian?

```{r}
beeswax <- read.csv("http://www3.amherst.edu/~nhorton/rice/chapter10/beeswax.csv")
hcarb <- beeswax$Hydrocarbon

hcarb.ecdf <- ecdf(hcarb)
plot(hcarb.ecdf)
histogram(hcarb)
```

```{r, fig.keep='last'}
qqnorm(hcarb)
qqline(hcarb)
```

```{r}
quantile(hcarb, c(.90, .75, .50, .25, .10))
```

The distribution does appear Gaussian.

## Part (b)
The average percentage of hydrocarbons in microcrystalline wax (a synthetic commercial wax) is 85%. Suppose that beeswax was diluted with 1% microcrystalline wax. Could this be detected? What about a 3% or a 5% dilution?

To help illustrate these dilutions, we plot histograms of the original beeswax compared to histograms of the diluted beeswaxes.

```{r, fig.keep='none'}
undiluted <- hist(hcarb)
synth.01 <- hist(hcarb*.99 + 85*.01)
synth.03 <- hist(hcarb*.97 + 85*.03)
synth.05 <- hist(hcarb*.95 + 85*.05)
```
```{r, fig.keep='last'}
plot(undiluted, col=rgb(0,0,1,1/4), xlim=c(11, 21), ylim = c(0, 20))
plot(synth.01, col=rgb(1,0,0,1/4), xlim=c(11, 21), add = TRUE)
(mean(hcarb*.99 + 85*.01) - mean(hcarb))
```

If we examine the mean of the 1% dilutions, we notice that it is only .70 standard deviations away from the mean of the originals, i.e. within 1 standard deviation of the original mean. As the histograms above show, it would be difficult to detect a 1% dilution. 

```{r, fig.keep='last'}
plot(undiluted, col=rgb(0,0,1,1/4), xlim=c(11, 21), ylim = c(0, 20))
plot(synth.03, col=rgb(0,1,0,1/4), xlim=c(11, 21), add = TRUE)
(mean(hcarb*.97 + 85*.03) - mean(hcarb))
```

If we examine the mean of the 3% dilutions, we notice that it is 2.11 standard deviations away from the mean of the originals, which is not much more than 2 standard deviations. As the histograms above show, it would be somewhat difficult to detect a 3% dilution.

```{r, fig.keep='last'}
plot(undiluted, col=rgb(0,0,1,1/4), xlim=c(11, 21), ylim = c(0, 20))
plot(synth.05, col=rgb(1,1,0,1/4), xlim=c(11, 21), add = TRUE)
(mean(hcarb*.95 + 85*.05) - mean(hcarb))
```

If we examine the mean of the 5% dilutions, we notice that it is 3.52 standard deviations away from the mean of the originals, which is more than 3 standard deviations away from the original mean. As the histograms above show, it would be relatively easy to detect a 5% dilution.

# Problem 10.10

Let's take a sample of size $n$ from a Normal distribution with mean $\mu$ and standard deviation $\sigma$. We want to determine the variance of the sample median and compare it to the variance of the sample mean.

To make life easy, let's use $n = 99$, $\mu = 0$, and $\sigma = 1$. We will simulate draws of size 99 from a Normal(0,1) and examine the sample medians and means (This is part (d)).

```{r}
numsim <- 10000

n <- 99 # Number of elements in each sample
mean <- 0
sd <- 1

medVals <- do(numsim)*median(rnorm(n, mean = mean, sd = sd))
var(medVals[,1])
```

So, the variance of the sample median is `r var(medVals[,1])`.

```{r}
meanVals <- do(numsim)*mean(rnorm(n, mean = mean, sd = sd))
var(meanVals[,1])
```

So, the variance of the sample means is `r var(meanVals[,1])`.

If we look at the ratio between the variances, we get `r var(medVals[,1])/var(meanVals[,1])`, which looks suspiciously like $\frac{\pi}{2} \approx 1.571$. In fact, it is $\frac{\pi}{2}$, as we will show with the following (heavily simplified) analytical excursion.

To find the analytical ratio between the variance of the sample median and the sample mean, we first need to find the actual variances of the sample median and the sample mean. Rice gives us a formula for the approximate variance of the kth order statistic (see part (b)). If we plug in k = 50 for the median of a sample of n = 99, we get:

$$Var(Y_{50}) \approx \frac{1}{2}(1-\frac{1}{2}) (\frac{1}{f (F^{-1}(\frac{1}{2}))^{2}}) (\frac{1}{n})$$

Because we are using a Normal(0,1), we know that $F^{-1}(1/2)$ is just the mean, which is just the median (See part (a) for more generality). So, we can just substitute in $\mu = 0$.

$$Var(Y_{50}) \approx \frac{1}{2}(1-\frac{1}{2}) (\frac{1}{f (0)^{2}}) (\frac{1}{n})$$

Now, we evaluate the pdf of the Normal(0,1) at 0.

$$f_{x}(x) = \frac{1}{\sqrt{2 \pi \sigma}} exp[\frac{-(x-\mu)^{2}}{2\sigma^{2}}]$$
$$f_{x}(0) = \frac{1}{\sqrt{2 \pi (1)}} exp[\frac{-(0-0)^{2}}{2(1)}] = \frac{1}{\sqrt{2 \pi}}$$

Plugging back in, we get:

$$Var(Y_{50}) \approx \frac{1}{2}(1-\frac{1}{2}) (\frac{1}{(\frac{1}{\sqrt{2 \pi})^{2}}}) (\frac{1}{101})$$
$$Var(Y_{50}) \approx \frac{1}{2}(\frac{1}{2})(2 \pi) (\frac{1}{n})$$

So, we have found the variance of the median to be $\frac{\pi}{2n}$.

From Rice (page 207), we know that the variance of the sample mean is $\frac{\sigma^2}{n}$, which for us is just $\frac{1}{n}$. Thus, the ratio of the variance of the sample median to the variance of the sample mean is just $\frac{\pi}{2}$.

# Problem 10.28

For a sample of size n = 3, median = $\eta$, we know that:
$$P(x_{1} < \eta < x_{2}) = 1 - P(\eta < x_{1}) - P(\eta > x_{2})$$
$$ = 1 - P(0~observations~less~than~\eta) - P(0~observations~greater~than~\eta) - P(1~observation~greater~than~\eta)$$

Using binomial distributions with p = .5 (since the probability that any random observation is greater than or less than the median is .5), we get:
$$P(x_{1} < \eta < x_{2})  = 1 - \frac{1}{8} - \frac{1}{8} - \frac{3}{8} = \frac{3}{8}$$

Empirically, we use a standard Normal as our continuous probability distribution. So, the median = 0. 
```{r}
isBetween12 <- function() {
  sample <- rnorm(3, 0, 1)
  order1 <- min(sample)
  order2 <- median(sample) 
  return(order1<0 & order2>0)
}
x <- do(10000) * isBetween12()
tally(~x, format='prop')
```

We get an empirical solution of about $\frac{3}{8}$.

For the second part:
$$P(x_{1} < \eta < x_{3}) = 1 - P(\eta < x_{1}) - P(\eta > x_{3})$$
$$ = 1 - P(0~observations~less~than~\eta) - P(0~observations~greater~than~\eta)$$

Again using binomial distributions with p = .5, we get:
$$P(x_{1} < \eta < x_{2})  = 1 - \frac{1}{8} - \frac{1}{8} = \frac{6}{8} = \frac{3}{4}$$

Empirically, we use a standard Normal again (median = 0).
```{r}
isBetween13 <- function() { 
  sample <- rnorm(3, 0, 1)
  order1 <- min(sample)
  order3 <- max(sample)
  return(order1<0 & order3>0)
}
x <- do(10000) * isBetween13()
tally(~x, format='prop')
```
We get an empirical solution of about $\frac{3}{4}$.